---
title: "Predict431-TS9-CODE"

output: word_document
    
---
<style>
body {
    position: absolute;
    left: 10px;}
</style>

<style type="text/css">

.main-container {
  max-width: 1280px;
}

</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning =FALSE, error=FALSE,message=FALSE)


```



```{r}

setwd("C:/Users/Admin/Dropbox/Northwestern University/Predict413/Final/")

```


```{r}

# install.packages("PerformanceAnalytics")
#### load libraries #### 

library(PerformanceAnalytics) 

library(quantmod)
library(rugarch)
library(car)
library(FinTS)
library(rmgarch)
options(digits=4)

#### import data #### 

# download data 
symbol.vec = c("XOM", "CVX")
getSymbols(symbol.vec, from ="2013-01-01", to = "2016-12-31") 


# View(XOM)
# View(CVX)

```



```{r}

options(digits=8)
library("fBasics")
basicStats(XOM)
basicStats(CVX)

```



```{r}

par(mfcol=c(2,2))

h5<-hist(XOM$XOM.Adjusted , breaks=20, main="XOM", xlab="Adjusted Prices ", col = "blue")

h6<-hist(CVX$CVX.Adjusted , breaks=20, main="CVX", xlab="AdjustedPrices" , col = "green")



qqnorm(XOM$XOM.Adjusted, main = "Q-Q Plot for XOM",
             col =  "blue",cex.axis = 1.5, cex = 1.5); qqline(XOM$XOM.Adjusted)

qqnorm(CVX$CVX.Adjusted, main = "Q-Q Plot for CVX",
             col =  "green",cex.axis = 1.5, cex = 1.5); qqline(CVX$CVX.Adjusted)

par(mfcol=c(1,1))

```
```{r}

# plot prices

plot(XOM$XOM.Adjusted, main ="XOM Daily Adjusted Price ")
plot(CVX$CVX.Adjusted, main ="CVX Daily Adjusted Price ")


```


```{r}

# calculate log-returns for GARCH analysis
XOM.ret = CalculateReturns(XOM$XOM.Adjusted, method="log")
CVX.ret = CalculateReturns(CVX$CVX.Adjusted, method="log")

XOM.ret<- XOM.ret[-1]
CVX.ret<- CVX.ret[-1]

colnames(XOM.ret) ="XOM"
colnames(CVX.ret) = "CVX"

```


```{r}

par(mfcol=c(2,2))

h5<-hist(XOM.ret, breaks=20, main="XOM.Adjusted ", xlab="log-returns" , col = "blue")

h6<-hist(CVX.ret, breaks=20, main="CVX.Adjusted", xlab="log-returns", col = "green")

qqnorm(XOM.ret, main = "Q-Q Plot for XOM log-returns",
             col =  "blue",cex.axis = 1.5, cex = 1.5); qqline(XOM.ret)
qqnorm(CVX.ret, main = "Q-Q Plot for CVX log-returns",
             col =  "green",cex.axis = 1.5, cex = 1.5); qqline(CVX.ret)

par(mfcol=c(1,1))

```


```{r}

# create combined data series
XOM.CVX.ret = merge(XOM.ret,CVX.ret)

```



```{r}

basicStats_combined <- basicStats(XOM.CVX.ret)

library("knitr","xtable")

kable(basicStats_combined[c('Mean', 'Median', 'Stdev', 'Skewness', 'Kurtosis', 'Minimum', 'Maximum'), ],
      format="pandoc", caption="Basic Statistics of XOM and CVX log returns")

```





```{r}

# plot returns
plot(XOM.ret)
plot(CVX.ret)


```


```{r}

# scatterplot of returns
plot( coredata(XOM.ret), coredata(CVX.ret), xlab="XOM log return", ylab="CVX log return",
      type="p", pch=16, lwd=2, col="blue")
abline(h=0,v=0)


```




```{r}


d1=density(XOM.ret)
d2=density(CVX.ret)

par(mfcol=c(1,2))
plot(d1$x,d1$y,xlab='returns',ylab='density',main="XOM Log Returns Density",type='l')
plot(d2$x,d2$y,xlab='returns',ylab='density',main="CVX Log Returns Density",type='l')



```


```{r}

# Shapiro-Wilk test
# shapiro.test() strictly takes in vectors.

shapiro.test(as.vector(XOM.ret))

shapiro.test(as.vector(CVX.ret))


```






### 1.2 Tests

Include a test of the null hypothesis that the mean of each of the series returns is zero. 


```{r}
# XOM

t.test(XOM.ret)



# CVX

t.test(CVX.ret)

```


Third Moment: Skewness

Test H0: m3 = 0 vs. Ha: m3 != 0, where m3 denotes skewness of the log returns

Testing the symmetry of the series with respect to the mean

Compute p-value


```{r}

# XOM

XOM.ret.skew <- skewness(XOM.ret)/sqrt(6/length(XOM.ret))

XOM.ret.skew

XOM.ret.skew.pv <- 2*(1-pnorm(abs(XOM.ret.skew)))

XOM.ret.skew.pv


# CVX

CVX.ret.skew <- skewness(CVX.ret)/sqrt(6/length(CVX.ret))

CVX.ret.skew

CVX.ret.skew.pv <- 2*(1-pnorm(abs(CVX.ret.skew)))

CVX.ret.skew.pv


```

 Test H0 : K = 3 vs. Ha : K 6= 3, where K denotes the kurtosis.
 
 
```{r}


# XOM.ret

XOM.ret.kurt <- kurtosis(XOM.ret)/sqrt(24/length(XOM.ret))

XOM.ret.kurt

XOM.ret.kurt.pv <- 2*(1-pnorm(abs(XOM.ret.kurt))); XOM.ret.kurt.pv



# CVX

CVX.ret.kurt <- kurtosis(CVX.ret)/sqrt(24/length(CVX.ret))
CVX.ret.kurt
CVX.ret.kurt.pv <- 2*(1-pnorm(abs(CVX.ret.kurt))); CVX.ret.kurt.pv


```




### 1.3 Box-Jenkins

Use the Box-Jenkins methodology to perform univariate time series model ???tting to each of the series. Include details of each step of the model-building process, and support your ???nal model selection for each series.


-- The Box-Jenkins methodology is a strategy or procedure that can be used to build an ARIMA model. 

if you keep repeating this process - estimating lots and lots of ARIMA models - you will eventually be able to select a model with the lowest Akaike Information criterion (from the set of models that you have estimated). In this context, a more systematic approach would be to use an algorithm based on comparing AIC values for a variety of different models to automatically select an ARIMA model for you, such as the one provided by the forecast package in R - the relevant function name is auto.arima().

In any event, the procedure you outlined involved selecting an ARIMA model based on minimizing some information criterion (in this case, AIC, but there are other measures




```{r}


# Model Identification

#--------------------------------------



# Apply ADF test - check for unit-root nonstationarity

# If p-value > alpha, then fail to reject H0: that series is unit-root

#   nonstationary (i.e. a unit-root is likely present); suggests differencing

#   is required


library("tseries")

adf.test(XOM.ret, alternative = "stationary")


adf.test(CVX.ret, alternative = "stationary")




```


```{r}
#library ("forecast")
#tsdisplay(XOM.ret, main = "XOM.ret")


# Are there serial correlations in the log return?

# ACF plot of series for checking serial correlations in series

# ACF plot of abs(series) for checking dependence in series

library ("forecast")
tsdisplay(XOM.ret, main = "XOM Daily Log Return")

#acf(XOM.ret, 25, xlim = c(1, 25), ylim = c(-0.2, 0.2))
#pacf(XOM.ret, 25, ylim = c(-0.1, 0.1))


tsdisplay(CVX.ret, main = " CVX Daily Log Return")

```
```{r}

m1 = ar(XOM.ret)
print(m1)


m2 = ar(CVX.ret)
print(m2)

```
```{r}

auto.arima(XOM.ret, stationary = T, trace = T, approx = F, allowmean = F)


auto.arima(XOM.ret, seasonal=FALSE)


```

```{r}

adf.test(XOM.ret, k=4)

```


```{r}

Box.test(XOM.ret, lag = 4, type = "Ljung")


Box.test(CVX.ret, lag = 10, type = "Ljung")

```




We'll now use the arima method to create an AR(4,0,0) model for XOM.ret

```{r}

m1_xom = arima(XOM.ret, order = c(4, 0, 0), include.mean = FALSE)
print(m1_xom)

```



```{r  fig.width=10,fig.height=10}

tsdiag(m1_xom, gof.lag = 24)

```

```{r}

Box.test(m1_xom$residuals, lag = 10, type = "Ljung")

```

```{r}

acf(residuals(m1_xom), 25, xlim = c(1, 25), ylim = c(-0.2, 0.2))

pacf(residuals(m1_xom), 25, ylim = c(-0.2, 0.2))

```

```{r}



hist(m1_xom$residuals , breaks=20, main="XOM Model Residuals" , col = "blue")



qqnorm(m1_xom$residuals, main = "Q-Q Plot for XOM Model Residuals",
             col =  "blue",cex.axis = 1.5, cex = 1.5); qqline(XOM.ret)

par(mfcol=c(1,1))

```


```{r}

auto.arima(CVX.ret, stationary = T, trace = T, approx = F, allowmean = F)

```

```{r}

m1_cvx = arima(CVX.ret, order = c(2, 0, 2), include.mean = FALSE)
print(m1_cvx)


```
```{r  fig.width=10,fig.height=10}

tsdiag(m1_cvx, gof.lag = 24)

```


```{r}

Box.test(m1_cvx$residuals, lag = 10, type = "Ljung")

```

```{r}

acf(residuals(m1_cvx), 25, xlim = c(1, 25), ylim = c(-0.2, 0.2))

pacf(residuals(m1_cvx), 25, ylim = c(-0.2, 0.2))

```



```{r}



hist(m1_cvx$residuals , breaks=20, main="CVX Model Residuals" , col = "green")



qqnorm(m1_cvx$residuals, main = "Q-Q Plot for CVX Model Residuals",
             col =  "green",cex.axis = 1.5, cex = 1.5); qqline(XOM.ret)

par(mfcol=c(1,1))

```


#### 1.3 Forecast

```{r}

library("forecast")


xom_forecast <- forecast.Arima(m1_xom, h=19)
xom_forecast

```
```{r}

plot.forecast(xom_forecast)

```

```{r}



cvx_forecast <- forecast.Arima(m1_cvx, h=19)
cvx_forecast


```


```{r}

plot.forecast(cvx_forecast)

```
```{r}

# Are there ARCH effects in the log return series? Why or why not?

# Ljung-Box test

#------------------

# Use squared series

# H0: first m lags of ACF of squared series = 0

# If we reject H0:, series shows strong ARCH effects



# Test at lag 10 - reject H0:

XOM.ret.arch <- (XOM.ret - mean(XOM.ret))

Box.test(XOM.ret.arch^2, lag = 10, type = "Ljung")

```
```{r}


acf(XOM.ret.arch^2, 25, xlim = c(1, 25), ylim = c(-0.2, 0.2))

pacf(XOM.ret.arch^2, 25, ylim = c(-0.2, 0.2))
     
     
```



```{r}


library(fGarch)
XOM.m2 <- garchFit(~arma(4,0)+garch(1, 1), data = XOM.ret.arch, trace = F)

summary(XOM.m2)



```



```{r}

XOM.m2.res <- residuals(XOM.m2, standardize = T)



# Q-Q Plot

qqnorm(XOM.m2.res); qqline(XOM.m2.res)

```

```{r}


# Shapiro test of normality - H0: iid normal

shapiro.test(XOM.m2.res)

```

```{r}

# Model 2: GARCH(1,1) | Student-t Innovation

# Build model

XOM.m3 <- garchFit(~arma(4, 0) +garch(1, 1), data = XOM.ret.arch, trace = F, 
                       cond.dist = "std" , include.mean = F)


summary(XOM.m3)




```



```{r}


XOM.m3.res <- residuals(XOM.m3, standardize = T)



# Q-Q Plot

qqnorm(XOM.m3.res); qqline(XOM.m3.res)


```


```{r}



# Shapiro test of normality - H0: iid normal

shapiro.test(XOM.m3.res)


```


```{r}

acf(XOM.m3.res, 25, xlim = c(1, 25), ylim = c(-0.1, 0.1))

pacf(XOM.m3.res, 25, ylim = c(-0.1, 0.1))

```





```{r}


CVX.ret.arch <- (CVX.ret - mean(CVX.ret))

Box.test(CVX.ret.arch^2, lag = 10, type = "Ljung")

```

```{r}


acf(CVX.ret.arch^2, 25, xlim = c(1, 25), ylim = c(-0.2, 0.2))


pacf(CVX.ret.arch^2, 25, ylim = c(-0.2, 0.2))
     
     
```


```{r}


library(fGarch)

CVX.m2 <- garchFit(~arma(2,2)+garch(1, 1), data = CVX.ret.arch, trace = F)

summary(CVX.m2)

```

```{r}

CVX.m2.res <- residuals(CVX.m2, standardize = T)

# Q-Q Plot

qqnorm(CVX.m2.res); qqline(CVX.m2.res)
```

```{r}

# Shapiro test of normality - H0: iid normal

shapiro.test(CVX.m2.res)

```

```{r}

acf(CVX.m2.res, 25, xlim = c(1, 25), ylim = c(-0.1, 0.1))

pacf(CVX.m2.res, 25, ylim = c(-0.1, 0.1))

```

```{r}

# Model 2: GARCH(1,1) | Student-t Innovation

# Build model

CVX.m3 <- garchFit(~arma(2,2)+garch(1, 1), data = CVX.ret.arch, trace = F, 
                       cond.dist = "std", include.mean = F)

summary(CVX.m3)
```


```{r}

CVX.m3.res <- residuals(CVX.m3, standardize = T)

# Q-Q Plot

qqnorm(CVX.m3.res); qqline(CVX.m3.res)

```

```{r}

shapiro.test(CVX.m3.res)

```

```{r}

acf(CVX.m3.res, 25, xlim = c(1, 25), ylim = c(-0.1, 0.1))

pacf(CVX.m3.res, 25, ylim = c(-0.1, 0.1))


```

```{r}


# Obtain the 1-step to 5-step ahead mean and volatility forecasts using the

#   fitted ARMA-GARCH model with Student-t innovations.



# Note: standardDeviation value is volatility forecast

xom.predict <- predict(XOM.m3, n.ahead = 1, h=19); xom.predict



```

```{r}

cvx.predict <- predict(CVX.m3, n.ahead = 1); xom.predict

```


```{r}

"ccm" <- function(x,lags=12,level=FALSE){
# Compute and plot the cross-correlation matrices.
# lags: number of lags used.
# level: logical unit for printing.
# Created by R.S. Tsay, March 2009
#
if(!is.matrix(x))x=as.matrix(x)
T=nrow(x)
k=ncol(x)
if(lags < 1)lags=1

# remove the sample means
av=apply(x,2,mean)
y=x
for (j in 1:k){
y[,j]=x[,j]-av[j]
}
V1=cov(y)
print("Covariance matrix:")
print(V1,digits=3)
se=sqrt(diag(V1))
SD=diag(1/se)
S0=SD%*%V1%*%SD
ksq=k*k
wk=matrix(0,ksq,(lags+1))
wk[,1]=c(S0)
j=0
cat("CCM at lag: ",j,"\n")
print(S0,digits=3)
y=y%*%SD
crit=2.326/sqrt(T)
cat("Simplified matrix:","\n")
for (j in 1:lags){
y1=y[1:(T-j),]
y2=y[(j+1):T,]
Sj=t(y2)%*%y1/T
Smtx=matrix(".",k,k)
for (ii in 1:k){
for (jj in 1:k){
if(Sj[ii,jj] > crit)Smtx[ii,jj]="+"
if(Sj[ii,jj] < -crit)Smtx[ii,jj]="-"
}
}
cat("CCM at lag: ",j,"\n")
for (ii in 1:k){
cat(Smtx[ii,],"\n")
}
if(level){
cat("Correlations:","\n")
print(Sj,digits=3)
}
wk[,(j+1)]=c(Sj)

}
par(mfcol=c(k,k))
if(k > 3)par(mfcol=c(3,3))
tdx=c(0,1:lags)
jcnt=0
for (j in 1:ksq){
plot(tdx,wk[j,],type='h',xlab='lag',ylab='ccf',ylim=c(-1,1))
abline(h=c(0))
crit=2/sqrt(T)
abline(h=c(crit),lty=2)
abline(h=c(-crit),lty=2)
jcnt=jcnt+1
if(jcnt==9){
jcnt=0
cat("Hit return for more: ","\n")
readline()
}
#
}
par(mfcol=c(1,1))

ccm<-list(ccm=wk)

}

```


```{r  fig.width=10,fig.height=10}

ccm(XOM.CVX.ret, 5)

```



```{r}

"mq" <- function(x,lag=24){
# Compute multivariate Ljung-Box test statistics
#
if(!is.matrix(x))x=as.matrix(x)
nr=nrow(x)
nc=ncol(x)
g0=var(x)
ginv=solve(g0)
qm=0.0
cat("        m,  Q(m)    and p-value:","\n")
df = 0
pvs=NULL
for (i in 1:lag){
  x1=x[(i+1):nr,]
  x2=x[1:(nr-i),]
  g = cov(x1,x2)
  g = g*(nr-i-1)/(nr-1)
  h=t(g)%*%ginv%*%g%*%ginv
  qm=qm+nr*nr*sum(diag(h))/(nr-i)
  df=df+nc*nc
  pv=1-pchisq(qm,df)
  pvs=c(pvs,pv)
  print(c(i,qm,pv),digits=5)
}
par(mfcol=c(1,1))
plot(pvs,ylim=c(0,1),main="p-values of Ljung-Box statistics")
abline(h=c(0))
lines(rep(0.05,lag),lty=2,col='blue')
}

```

```{r}

# Test the following hypotheses using a 5% significance level:

#   H0: p1 = ... = pn = 0

#   Ha: pi != 0 for some i, where {1, ..., n}

mq(XOM.CVX.ret, lag = 10)

```
```{r}
# Using a 30-day moving window, compute and plot rolling covariances and

#   correlations. Briefly comment on what you see.



#--------------------------------------

# Compute and plot rolling correlations and covariances

#--------------------------------------



# Specify correlation & covariance functions

cor.fun <- function(x){

    cor(x)[1,2]

}



cov.fun <- function(x){

    cov(x)[1,2]

}



# Window: 30 days

returns.roll.cov <- rollapply(as.zoo(XOM.CVX.ret), FUN = cov.fun, width = 30,

                              by.column = F, align = "right")

returns.roll.cor <- rollapply(as.zoo(XOM.CVX.ret), FUN = cor.fun, width = 30,

                              by.column = F, align = "right")

# Plots

plot(returns.roll.cov, main = "XOM & CVX: 30-day Rolling Covariances",

     ylab = "Covariance", lwd = 2, col = "blue")

grid()

abline(h = cov(XOM.CVX.ret)[1,2], lwd = 2, col = "red")



plot(returns.roll.cor, main = "XOM & CVX: 30-day Rolling Correlations",

     ylab = "Correlation", lwd = 2, col = "blue")

grid()

abline(h = cor(XOM.CVX.ret)[1,2], lwd = 2, col = "red")


```

```{r}

#--------------------------------------

# Compute and plot EWMA correlations and covariances

# Use lambda = 0.94

#--------------------------------------



# EWMA covariance

returns.cov.ewma <- covEWMA(as.data.frame(XOM.CVX.ret), lambda = 0.94)



# EWMA conditional covariance

returns.cond.cov <- returns.cov.ewma[, 2, 1]



# EWMA conditional correlation

t <- length(returns.cov.ewma[, 1, 1])

returns.cond.cor <- rep(0, t)

for (i in 1:t) {

    returns.cond.cor[i]<- cov2cor(returns.cov.ewma[i, , ])[1, 2]

}



# Plots

plot(x = time(as.zoo(XOM.CVX.ret)), y = returns.cond.cov, type = "l", xlab = "Time", 

     ylab = "Covariance", lwd = 2, col = "blue",

     main = "XOM & CVX: EWMA Covariance")

grid()

abline(h = cov(XOM.CVX.ret)[1, 2], lwd = 2, col = "red")



plot(x = time(as.zoo(XOM.CVX.ret)), y = returns.cond.cor, type = "l", xlab = "Time",

     ylab = "Correlation", lwd = 2, col = "blue",

     main = "XOM & CVX: EWMA Correlation")

grid()

abline(h = cor(XOM.CVX.ret)[1, 2], lwd = 2, col = "red")


```

```{r}

# Use dccfit() from {rmgarch}, and estimate the normal-DCC(1,1) model

# Comment on the estimated coefficients and the model fit



# Specification of DCC model for each series - univariate normal GARCH(1,1)
#
# DCC estimation
#

# univariate normal GARCH(1,1) for each series
garch11.spec = ugarchspec(mean.model = list(armaOrder = c(0,0)), 
                          variance.model = list(garchOrder = c(1,1), 
                          model = "sGARCH"), 
                          distribution.model = "norm")

# dcc specification - GARCH(1,1) for conditional correlations

# Multivariate specification of DCC specification for conditional correlations
dcc.garch11.spec = dccspec(uspec = multispec( replicate(2, garch11.spec) ), 
                           dccOrder = c(1,1), 
                           distribution = "mvnorm")

# Validate specification of DCC model
dcc.garch11.spec

```
```{r}

# Build model

dcc.fit = dccfit(dcc.garch11.spec, data = XOM.CVX.ret)

# Explore available parameters and results
class(dcc.fit)
slotNames(dcc.fit)
names(dcc.fit@mfit)
names(dcc.fit@model)

# many extractor functions - see help on DCCfit object
# coef, likelihood, rshape, rskew, fitted, sigma, 
# residuals, plot, infocriteria, rcor, rcov
# show, nisurface

# show dcc fit
dcc.fit
```
```{r}


# Examine residuals for normality assumption

# Assign standardized residuals




hist(dcc.fit@mfit$stdresid , breaks=20, main="dcc.fit residuals", xlab="Model Residuals")


# Q-Q Plot

qqnorm(dcc.fit@mfit$stdresid); qqline(dcc.fit@mfit$stdresid)

```


```{r}

"mq" <- function(x,lag=24){
# Compute multivariate Ljung-Box test statistics
#
if(!is.matrix(x))x=as.matrix(x)
nr=nrow(x)
nc=ncol(x)
g0=var(x)
ginv=solve(g0)
qm=0.0
cat("        m,  Q(m)    and p-value:","\n")
df = 0
pvs=NULL
for (i in 1:lag){
  x1=x[(i+1):nr,]
  x2=x[1:(nr-i),]
  g = cov(x1,x2)
  g = g*(nr-i-1)/(nr-1)
  h=t(g)%*%ginv%*%g%*%ginv
  qm=qm+nr*nr*sum(diag(h))/(nr-i)
  df=df+nc*nc
  pv=1-pchisq(qm,df)
  pvs=c(pvs,pv)
  print(c(i,qm,pv),digits=5)
}
par(mfcol=c(1,1))
plot(pvs,ylim=c(0,1),main="p-values of Ljung-Box statistics")
abline(h=c(0))
lines(rep(0.05,lag),lty=2,col='blue')
}

```

```{r}


# Test the following hypotheses using a 5% significance level:

#   H0: p1 = ... = pn = 0

#   Ha: pi != 0 for some i, where {1, ..., n}

mq(dcc.fit@mfit$stdresid)

```




```{r  fig.width=10,fig.height=10}
# conditional sd of each series
plot(dcc.fit, which=2)

```


```{r}
# conditional correlation
plot(dcc.fit, which=4, main="", sub="")


```

```{r}
# extracting correlation series
ts.plot(rcor(dcc.fit)[1,2,])

```



```{r}
# Q1F

#======================================



# Compute and plot the first 100 h-step ahead forecasts of the conditional

#   covariance and correlation of the fitted model


#
# forecasting conditional volatility and correlations
#

# Build model

dcc.fcst = dccforecast(dcc.fit, n.ahead=19)

# Explore available parameters and results

class(dcc.fcst)
slotNames(dcc.fcst)
class(dcc.fcst@mforecast)
names(dcc.fcst@mforecast)

# many method functions - see help on DCCforecast class
# rshape, rskew, fitted, sigma, plot, rcor, rcov, show


```


```{r}


# show forecasts
dcc.fcst

```


```{r}

# Method 1

# Note: second part fails, see Method 2 below

#------------------

# Conditional Covariances

plot(dcc.fcst, which = 3)

```