
---
title: "Predict431-TS8-CODE"
output:
  html_document: default
  word_document: default
---
<style>
body {
    position: absolute;
    left: 10px;}
</style>

<style type="text/css">

.main-container {
  max-width: 1280px;
}

</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning =FALSE, error=FALSE,message=FALSE)


```




```{r}

setwd("C:/Users/Admin/Dropbox/Northwestern University/Predict413/Assignment 8/")

```


1.1. Use EDA to justify a log transformation and a ???rst di???erence transformation, zit, of each time series for i = 1,2 hb???n and hbfrbn, respectively.

```{r}

require(MTS)

da1 <- read.table("q-fdebt.txt",header=T)

# str(da1)

fin <-da1$hbfin

frbn <-da1$hbfrbn

```

#### Q1:EDA


Data Summaries and Tables


```{r}

library("fBasics")
basicStats_da1 <- basicStats(da1)

library("knitr","xtable")

kable(basicStats_da1[c('Mean', 'Median', 'Stdev', 'Skewness', 'Kurtosis', 'Minimum', 'Maximum'), c(3:4)],
      format="pandoc", caption="Basic Statistics of U.S. quarterly federal debts")


```

```{r}

par(mfcol=c(2,2))

h1<-hist(fin, breaks=20, main="Debt Held by\n Foreign and International Investors ", xlab=" Debt")
# axis(side=1, at=seq(0,82,1), pos=0, las=0)

abline(v = mean(fin , na.rm = TRUE), col = "red", lwd = 2, lty = 2)
abline(v = median(fin, na.rm = TRUE), col = "green", lwd = 2, lty = 2)

xfit1<-seq(min(fin, na.rm = TRUE),max(fin, na.rm = TRUE), length=40)
yfit1<-dnorm(xfit1, mean = mean(fin, na.rm = TRUE), sd=sd(fin, na.rm = TRUE))
yfit1<-yfit1*diff(h1$mids[1:2]*length(fin))
lines(xfit1, yfit1, col="blue",lwd=2)

legend("topright",  # location of legend within plot area
c("Density plot", "Mean", "Median"),
col = c("blue", "red", "green"),
lwd = c(2, 2, 2), lty= c(1, 2, 2), cex = 0.6)


h2<-hist(frbn, breaks=20, main="Debt Held by\n Federal Reserve Banks", xlab=" Debt")
# axis(side=1, at=seq(0,82,1), pos=0, las=0)

abline(v = mean(frbn , na.rm = TRUE), col = "red", lwd = 2, lty = 2)
abline(v = median(frbn, na.rm = TRUE), col = "green", lwd = 2, lty = 2)

xfit2<-seq(min(frbn, na.rm = TRUE),max(frbn, na.rm = TRUE), length=40)
yfit2<-dnorm(xfit2, mean = mean(frbn, na.rm = TRUE), sd=sd(frbn, na.rm = TRUE))
yfit2<-yfit2*diff(h2$mids[1:2]*length(frbn))
lines(xfit2, yfit2, col="blue",lwd=2)

legend("topright",  # location of legend within plot area
c("Density plot", "Mean", "Median"),
col = c("blue", "red", "green"),
lwd = c(2, 2, 2), lty= c(1, 2, 2), cex = 0.6)



qqnorm(fin); qqline(fin)


qqnorm(frbn); qqline(frbn)


```





```{r}

# Transform to log returns


log_da1 <- data.frame(da1[, 1:2], apply(da1[, 3:4], 2, log1p))

library(MTS)


tdx=da1[,1]+da1[,2]/12
MTSplot(log_da1[,3:4],tdx)



log_fin <-log_da1$hbfin

log_frbn <-log_da1$hbfrbn



library("fBasics")
basicStats_logda1 <- basicStats(log_da1)

library("knitr","xtable")

kable(basicStats_logda1[c('Mean', 'Median', 'Stdev', 'Skewness', 'Kurtosis', 'Minimum', 'Maximum'), c(3:4)],
      format="pandoc", caption="Basic Statistics in Log")



```



```{r}

par(mfcol=c(2,2))

h3<-hist(log_fin, breaks=20, main="Log Debt Held by\n Foreign and International Investors ", xlab=" Debt")
# axis(side=1, at=seq(0,82,1), pos=0, las=0)

abline(v = mean(log_fin , na.rm = TRUE), col = "red", lwd = 2, lty = 2)
abline(v = median(log_fin, na.rm = TRUE), col = "green", lwd = 2, lty = 2)

xfit3<-seq(min(log_fin, na.rm = TRUE),max(log_fin, na.rm = TRUE), length=40)
yfit3<-dnorm(xfit3, mean = mean(log_fin, na.rm = TRUE), sd=sd(log_fin, na.rm = TRUE))
yfit3<-yfit3*diff(h3$mids[1:2]*length(log_fin))
lines(xfit3, yfit3, col="blue",lwd=2)

legend("topright",  # location of legend within plot area
c("Density plot", "Mean", "Median"),
col = c("blue", "red", "green"),
lwd = c(2, 2, 2), lty= c(1, 2, 2), cex = 0.6)


h4<-hist(log_frbn, breaks=20, main="Log Debt Held by\n Federal Reserve Banks", xlab=" Debt")
# axis(side=1, at=seq(0,82,1), pos=0, las=0)

abline(v = mean(log_frbn , na.rm = TRUE), col = "red", lwd = 2, lty = 2)
abline(v = median(log_frbn, na.rm = TRUE), col = "green", lwd = 2, lty = 2)

xfit4<-seq(min(log_frbn, na.rm = TRUE),max(log_frbn, na.rm = TRUE), length=40)
yfit4<-dnorm(xfit4, mean = mean(log_frbn, na.rm = TRUE), sd=sd(log_frbn, na.rm = TRUE))
yfit4<-yfit4*diff(h4$mids[1:2]*length(log_frbn))
lines(xfit4, yfit4, col="blue",lwd=2)

legend("topright",  # location of legend within plot area
c("Density plot", "Mean", "Median"),
col = c("blue", "red", "green"),
lwd = c(2, 2, 2), lty= c(1, 2, 2), cex = 0.6)


qqnorm(log_fin); qqline(log_fin)
qqnorm(log_frbn); qqline(log_frbn)


```


```{r}
# Assign first differenced series

z=apply(log_da1,2,diff)
z=data.frame(z[,3],z[,4])
colnames(z)=c("hbfin", "hbfrbn")


# Convert to time series

# ts.z <- ts(z, start = c(1970, 1), frequency = 4)



# Plot the time series




plot(z$hbfin, type="l",  xlab = "Year", ylab = "Debt Growth (log)", 

     main = "Differenced log hbfin")


plot(z$hbfrbn, type="l",  xlab = "Year", ylab = "Debt Growth (log)", 

     main = "Differenced log hbfrbn")


```



1.2. Obtain the ???rst 5 lags of sample cross-correlation matrices of the zit. 


```{r}

"ccm" <- function(x,lags=12,level=FALSE){
# Compute and plot the cross-correlation matrices.
# lags: number of lags used.
# level: logical unit for printing.
# Created by R.S. Tsay, March 2009
#
if(!is.matrix(x))x=as.matrix(x)
T=nrow(x)
k=ncol(x)
if(lags < 1)lags=1

# remove the sample means
av=apply(x,2,mean)
y=x
for (j in 1:k){
y[,j]=x[,j]-av[j]
}
V1=cov(y)
print("Covariance matrix:")
print(V1,digits=3)
se=sqrt(diag(V1))
SD=diag(1/se)
S0=SD%*%V1%*%SD
ksq=k*k
wk=matrix(0,ksq,(lags+1))
wk[,1]=c(S0)
j=0
cat("CCM at lag: ",j,"\n")
print(S0,digits=3)
y=y%*%SD
crit=2.326/sqrt(T)
cat("Simplified matrix:","\n")
for (j in 1:lags){
y1=y[1:(T-j),]
y2=y[(j+1):T,]
Sj=t(y2)%*%y1/T
Smtx=matrix(".",k,k)
for (ii in 1:k){
for (jj in 1:k){
if(Sj[ii,jj] > crit)Smtx[ii,jj]="+"
if(Sj[ii,jj] < -crit)Smtx[ii,jj]="-"
}
}
cat("CCM at lag: ",j,"\n")
for (ii in 1:k){
cat(Smtx[ii,],"\n")
}
if(level){
cat("Correlations:","\n")
print(Sj,digits=3)
}
wk[,(j+1)]=c(Sj)

}
par(mfcol=c(k,k))
if(k > 3)par(mfcol=c(3,3))
tdx=c(0,1:lags)
jcnt=0
for (j in 1:ksq){
plot(tdx,wk[j,],type='h',xlab='lag',ylab='ccf',ylim=c(-1,1))
abline(h=c(0))
crit=2/sqrt(T)
abline(h=c(crit),lty=2)
abline(h=c(-crit),lty=2)
jcnt=jcnt+1
if(jcnt==9){
jcnt=0
cat("Hit return for more: ","\n")
readline()
}
#
}
par(mfcol=c(1,1))

ccm<-list(ccm=wk)

}

```



```{r}

ccm(z,5)

```

1.3. Test H0 : ??1 = иии = ??10 = 0 versus Ha : ??j 6= 0 for some j, where j ???{1,...,10}. Draw the conclusion using the 5% signi???cance level.


```{r}

"mq" <- function(x,lag=24){
# Compute multivariate Ljung-Box test statistics
#
if(!is.matrix(x))x=as.matrix(x)
nr=nrow(x)
nc=ncol(x)
g0=var(x)
ginv=solve(g0)
qm=0.0
cat("        m,  Q(m)    and p-value:","\n")
df = 0
pvs=NULL
for (i in 1:lag){
  x1=x[(i+1):nr,]
  x2=x[1:(nr-i),]
  g = cov(x1,x2)
  g = g*(nr-i-1)/(nr-1)
  h=t(g)%*%ginv%*%g%*%ginv
  qm=qm+nr*nr*sum(diag(h))/(nr-i)
  df=df+nc*nc
  pv=1-pchisq(qm,df)
  pvs=c(pvs,pv)
  print(c(i,qm,pv),digits=5)
}
par(mfcol=c(1,1))
plot(pvs,ylim=c(0,1),main="p-values of Ljung-Box statistics")
abline(h=c(0))
lines(rep(0.05,lag),lty=2,col='blue')
}

```


```{r}

# Test the following hypotheses using a 5% significance level:

#   H0: p1 = ... = p10 = 0

#   Ha: pi != 0 for some i, where {1, ..., 10}

mq(z,10)

```

We can see all p-value is 0, so we can reject H0H0 of no cross-correlations with 5% significance level.

####----------------------------------------

2.1. Use EDA to justify a V AR(4) model

```{r}

data("mts-examples",package="MTS")
# head(qgdp)

# Plot the time series

library(MTS)


tdx=qgdp[,1]+qgdp[,2]/12
MTSplot(qgdp[,3:5],tdx)



```

#### Q1:EDA


Data Summaries and Tables


```{r}

library("fBasics")
basicStats_qgdp <- basicStats(qgdp)

library("knitr","xtable")
options(scipen=999)
kable(basicStats_qgdp[c('Mean', 'Median', 'Minimum', 'Maximum'), c(3:5)],
      format="pandoc", caption="Basic Statistics of basicStats_qgdp")


```


```{r}

par(mfcol=c(3,2))

h5<-hist(qgdp$uk, breaks=20, main="United Kingdom", xlab="GDP ")

h6<-hist(qgdp$ca, breaks=20, main="Canada", xlab="GDP ")

h7<-hist(qgdp$us, breaks=20, main="United States ", xlab="GDP ")

qqnorm(qgdp$uk); qqline(qgdp$uk)
qqnorm(qgdp$ca); qqline(qgdp$ca)
qqnorm(qgdp$us); qqline(qgdp$us)

```



```{r}


dat2=data.frame(qgdp$uk,qgdp$ca,qgdp$us)
colnames(dat2)=c("uk", "ca", "us")
logdat2=log(dat2)



library("fBasics")
basicStats_logdat2 <- basicStats(logdat2)

library("knitr","xtable")

kable(basicStats_logdat2[c('Mean', 'Median', 'Stdev', 'Skewness', 'Kurtosis', 'Minimum', 'Maximum'), ],
      format="pandoc", caption="Basic Statistics of log qgdp")




```




```{r}

par(mfcol=c(3,2))

h8<-hist(logdat2$uk, breaks=20, main="United Kingdom", xlab=" Log GDP ")

h9<-hist(logdat2$ca, breaks=20, main="Canada", xlab=" Log GDP ")

h10<-hist(logdat2$us, breaks=20, main="United States ", xlab="Log GDP ")

qqnorm(logdat2$uk); qqline(logdat2$uk)
qqnorm(logdat2$ca); qqline(logdat2$ca)
qqnorm(logdat2$us); qqline(logdat2$us)

```


```{r}


datgrowth=apply(logdat2,2,diff)
growth=100*datgrowth


# Determine Model Order




# Obtain the lags of the sample cross-correlation matricies of series

# Using level = T will output values and simplified notation

# ACFs are on primary diagonal and CCFs are on off diagonal

ccm(growth, 4)

```




```{r}

# Test the following hypotheses using a 5% significance level:

#   H0: p1 = ... = pn = 0

#   Ha: pi != 0 for some i, where {1, ..., n}

mq(growth)


```


```{r}

# Use VMAorder to determine model order

# Suggests VMA(2)

VMAorder(growth)

```


```{r}

# Apply t-test to check if H0: mean is not significantly different from zero

# Reject H0: in all series; include mean in modeled series

apply(growth, 2, t.test)

```


```{r}
# Model VAR(4)


# Build model
m1=VAR(growth,p=4)
m2=refVAR(m1, thres=1.96)
MTSdiag(m2)


```

